---
title: "Morning session"
format: html
editor: visual
execute:
  warning: TRUE
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(datapasta)
```

## Data wrangling refresher

1.  Starting from penguins, write a single piped seq in which:
    1.  only include penguins at dream and biscoe islands
    2.  remove the year and sex variables
    3.  add a new column body_mass_kg w penguin mass converted from g to kg
    4.  rename the island variable to location

```{r}
dream_biscoe <- penguins %>% 
  filter(island %in% c("Dream", "Biscoe")) %>% 
  select(-year, -sex) %>% 
  mutate(body_mass_kg = body_mass_g / 1000) %>% 
  rename(location = island)
```

2.  Write a single piped sequence to:
    2.  limit to Adelie penguins
    3.  Remove any observations where flipper_length_mm is NA (lis.na()) or drop_na)
    4.  Group the data by sex
    5.  Create a summary table that contains the mean, std, and sample size of flipper length for male and female Adelie penguins

```{r}
deli_sum <- penguins %>% 
  filter(species == "Adelie", !is.na(flipper_length_mm)) %>% 
  group_by(sex) %>% 
  summarise(mean_flipper = mean(flipper_length_mm), 
            sd(flipper_length_mm),
            sample_size = n())
```

```{r}
penguins %>% 
  group_by(species, island, sex) %>% 
  summarize(sample_size = n())

#Count is just wrapping together group_by + summarize  + n() to get counts of observations
penguins %>% 
  count(species, island, sex)
```

```{r}
#NOT REPRODUCABLE
animals <- data.frame(
  stringsAsFactors = FALSE,
          location = c("lagoon", "bluff", "creek", "oaks", "bluff"),
           species = c("bobcat", "coyote", "fox", "squirrel", "bobcat"),
          maturity = c("adult", "juvenile", "adult", "juvenile", "adult")
)
```

```{r}
sites <- data.frame(
  stringsAsFactors = FALSE,
          location = c("beach", "lagoon", "bluff", "oaks"),
    full_site_name = c("Goleta Beach","UCSB Lagoon",
                       "Ellwood Mesa","Fremont Campground"),
      jurisdiction = c("SB City", "UCSB", "SB City", "USFS")
)
```

```{r}
#Most conservative join
full_join_example <- full_join(animals, sites)
```

```{r}
#Left join example(keeps everything in left table, and only keep right table if it has exact match for key value)
left_join_example <- left_join(animals, sites)
```

```{r}
inner_join_examples <- inner_join(animals, sites)
```
